PInSoRo: A Dataset for Deep-learning of Human-Robot Social Interactions

### 2017-06-28 paper.tex
Update on Overleaf.

The acquisition of such a dataset requires first the definition of a suitable
task for the participants to perform while they are recorded. The nature of
this task influences in fundamental ways the kind of interactions that might be
observed, and thus, learned. Because they usually target very specific social
or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
usually simple, constrained tasks that do not reflect the complexity and dynamics of
real-world interactions. On the contrary, this dataset aims at capturing a rich set
of naturally-occurring social interactions. As such, a suitable task has to be:

\begin{itemize}
    \item elicit a large range of interaction situations;
    \item foster rich multi-modal interaction: simultaneous speech, gesture, and gaze
        behaviours are to be observed;
    \item fundamentally social, \ie the task would make little sense for an
        agent alone;
    \item foster non-trivial social dynamics, such as implicit turn-taking.
\end{itemize}

We present a new experimental task based on \emph{free play} interactions. Pairs
of children (4-7 years old) are invited to freely draw and interact with items
displayed on an interactive table, telling and acting stories with no explicit
goal set in advance (Fig.~\ref{fig|freeplay}). The task is designed so that
children can engage in open-ended and non-directive play situations, yet
sufficiently well defined to be controlled, suitable for detailled recording,
and easily reproducible.

@@

The \emph{freeplay sandbox} follows the sandtray
paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
with multitouch support) is used as an interactive surface (\emph{sandtray}). The two children play together
by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
depicts a generic empty environment, with different symbolic colours (water,
grass, beach, bushes...). By drawing on top of the background picture, the
children can also alter this environment to their liking. The players do not have any particular task to
complete, they are simply invited to freely play for as long as they wish.

In order to constraint the dataset to a tractable domain, the task is limited to
a dyadic \emph{face-to-face} interaction.  The resulting interaction space is
nevertheless high-dimensional, with two main axis: the game-related actions
(drawing, manipulating items, telling stories) and the social interactions
(monitoring the other's actions, discussing and agreeing on joint actions,
helping behaviours, etc.).

@@

The dataset consists in a collection of records. Each record correspond to one
play interaction between two children. To date (June 2017) 25 records have been acquired (\ie 50 children). At the end of the acquisition campaign
(July 2017), the dataset is planned to include 50 records. As
mentioned above, the duration of each play episode varies ($M=20m51s,
SD=10m40s$), but it is capped to a maximum of 40 minutes.

@@

\begin{table}[]
\centering
\caption{List of datastreams stored in each record}
\label{table|datastreams}
\begin{tabular}{@{}lll@{}}
\toprule
\bf Domain  & \bf Type                              & \bf Details                          \\ \midrule
child 1     & audio                                 & 16kHz, mono, semi-directional        \\
            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
            & facial features                       & 68 3D points (point cloud)           \\ \midrule
child 2     & audio                                 & 16kHz, mono, semi-directional        \\
            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
            & facial features                       & 68 3D points (point cloud)           \\ \midrule
environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
touchscreen & background drawing (RGB)              & 4Hz                                  \\
            & touches                               & 6 points multi-touch                 \\
            & items position and orientation        & ($x,y,\theta$)                       \\ \midrule
others      & \multicolumn{2}{l}{annotations of social behaviours and remarkable events}   \\
            & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
            & \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
\end{tabular}
\end{table}

@@

\subsection{Hardware and Software}

@@

While
the raw RGB and depth streams are neither rectified nor registered together (due
to performance considerations during the recording), it can be easily achieved
as a post-process step, using standard ROS tools (the ROS {\tt rgbd\_pipeline}).
Pre-configured scripts are available alongside the dataset\footnote{Available
online from
\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
republish registered streams and the corresponding 3D point-clouds (as seen in
Fig.~\ref{fig|rviz}, left).

>>>>

The acquisition of such a dataset requires first the definition of a suitable
task for the participants to perform while they are recorded. The nature of
this task influences in fundamental ways the kind of interactions that might be
observed, and later, learned. Because they typically target specific social
or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
usually simple, constrained tasks that do not reflect the complexity and dynamics of
real-world interactions. Conversely, this dataset aims at capturing a rich set
of naturally-occurring social interactions, and participants should engage in a task that:

\begin{itemize}
    \item is not directed, to maximise natural, non-contrived behaviours;
    \item elicit a large range of interaction situations;
    \item foster rich multi-modal interaction: simultaneous speech, gesture, and gaze
        behaviours are to be observed;
    \item evidence social dynamics, such as coupling, joint attention, turn-taking.
\end{itemize}

We have designed a new experimental task, called the \emph{freeplay sandbox}, and based on 
free play interactions. Pairs of children (4-8 years old) are invited to freely draw and interact with items
displayed on an interactive table, without any explicit goal set by the experimenter (Fig.~\ref{fig|freeplay}).
The task is designed so that children can engage in open-ended and non-directive play, yet it is
sufficiently defined to be suitable for recording, and reproducible in comparable conditions.

@@

The freeplay sandbox follows the sandtray
paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
with multitouch support) is used as an interactive surface (\emph{sandtray}). Two children play together
by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
depicts a generic empty environment, with different symbolic colours (water,
grass, beach, bushes...). By drawing on top of the background picture, the
children can change the environment to their liking. The players do not have any particular task to
complete, they are simply invited to freely play for as long as they want.

In order to constraint the dataset to a tractable domain, the task is limited to
a dyadic face-to-face interaction.  The resulting interaction space is
nevertheless high-dimensional, with two main axis: the game-related actions
(drawing, manipulating items, telling stories) and the social interactions
(monitoring the other's actions, discussing and agreeing on joint actions,
helping/hostile behaviours, etc.).

@@

The dataset consists in a collection of records. Each record correspond to one
play interaction between two children. To date (June 2017) 25 records have been
acquired (\ie 50 children), totalling 08h41m of recording. At the end of the acquisition campaign
(July 2017), the dataset is planned to include 50 records. As
mentioned above, the duration of each play episode varies ($M=20m51s,
SD=10m40s$), but it is capped to a maximum of 40 minutes.

@@

\begin{table}[]
\centering
\caption{List of datastreams stored in each record}
\label{table|datastreams}
\begin{tabular}{@{}lll@{}}
\toprule
\bf Domain  & \bf Type                              & \bf Details                          \\ \midrule
child 1     & audio                                 & 16kHz, mono, semi-directional        \\
            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
            & facial features                       & 68 3D points, 30Hz                   \\ \midrule
child 2     & audio                                 & 16kHz, mono, semi-directional        \\
            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
            & facial features                       & 68 3D points, 30Hz                   \\ \midrule
environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
touchscreen & background drawing (RGB)              & 4Hz                                  \\
            & touches                               & 6 points multi-touch, 10Hz           \\
            & items position and orientation        & ($x,y,\theta$), 10Hz                 \\ \midrule
annotations & \multicolumn{2}{l}{timestamped annotations of social behaviours and remarkable events}   \\ \midrule
other       & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
            & \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
\end{tabular}
\end{table}

@@

\subsection{Apparatus}

@@

The dataset contains the raw datastreams, as recorded during the study. No post-processing has been applied.
% in the robot condition, the interaction bag is trimmed to start once the robot is localised
While the raw RGB and depth video streams are neither rectified nor registered together (due
to performance considerations during the recording), it can be easily achieved
as a post-process step, using standard ROS tools (the ROS {\tt rgbd\_pipeline}).
Pre-configured scripts are available alongside the dataset\footnote{Available
online from
\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
republish registered streams and the corresponding 3D point-clouds (as seen in
Fig.~\ref{fig|rviz}, left).
